---
title: "Prediction Report"
author: "Chong, Chin Teck"
date: "January 31, 2016"
output: html_document
---

```{r global_option_downloading data, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```
#Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the [website here](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).


#Loading and Processing the Raw Data
The data for this project come from this [source](http://groupware.les.inf.puc-rio.br/har)

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

##Data downloading
We first set the default working directory and download the require training data and test data accordingly.

```{r Download data, echo=TRUE}
library(RCurl)
setwd("D:/project/Assignment/Assignment_8")

if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists("./data/pml-training.csv")) {
  url.training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  download.file(url.training, destfile = "./data/pml-training.csv")
}

if (!file.exists("./data/pml-testing.csv")) {
  url.testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
  download.file(url.training, destfile = "./data/pml-testing.csv")
}
```

##Reading data and data processing

First we would like to explore the data.  
```{R1, echo=TRUE}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```

Note that both dataset are having the same variables (160 variables). Next is try remove the near zero variance variables or columns that contain N/A missing values.
```{r remove N/A, echo=TRUE}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```
There were 107 variables with more than 95% of the data missing. Those variables were removed from the data as well. If we built a classification model based on those variables, then we can expect most of the time the variable is missing and therefore we cannot apply the classification rules on them. Therefore, building a model based on variables that's mostly missing is not practical.

##Data Partitioning
Partitioning Training data set into two data sets,70% for trainData, 30% for testData:

```{r data partitioning, echo=TRUE}
library(caret)
set.seed(12345) 
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

#Data Prediction and Modelling
Algorithm which will be used for the predictive model here is **Random Forest**

```{r prediction and modelling, echo=TRUE}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```

We estimate the performance of the model build on the validation data set. Getting the accuracy as well as the estimated out-of-sample error.
```{r estimate, echo=TRUE}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
accuracy <- postResample(predictRf, testData$classe)
error<-1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
```
The accuracy of the model is `accuracy *100`% and the estimated out-of-sample error is `error*100`%
